Parallel multi-objective optimization using co-evolutionary islands with overlapping search space

### 2017-02-10 hpmoon-coor.tex
Addressing more @JJ comments(not all of them). Also, a name to use instead cooperative-coevolutive is still needed

\begin{abstract}
  % First, state the problem!!!! - JJ
Multi-objective optimization methods have gained interest in the last
years because of the performance they can achieve solving real-world
problems. Many of these problems usually require a high-dimensional
decision space, and multi-objective Evolutionary Algorithms have been
successfully used in the past. Due to the large computational cost
these algorithms require, different parallelization and distribution
methods have been proposed, with different parallel and distributed
computer architectures to cope with them. One of the most extended
methods are the distributed co-evolutionary algorithms, where each
processor deals only with a part of the decision space, that is, the
individuals belonging to the same sub-population (island) explore the
same subset of decision variables. These subsets are usually disjoint
sets. We have previously proved that extending the length of these
subsets, and therefore, sharing some search space between islands,
implies an improvement in the results. In this paper, a new method to
automatically adapt the size of the overlapping, or shared, sections
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that coevolutionary
approaches can provide statistically significant improvements with
respect to the base algorithm, being the method to automatically adapt
the overlapping size the one to obtain the best performance during the
same execution time. Results also show that the relation of the number
of islands (subpopulations) to the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

%2 COMPUTATION -> PARALLEL
These kind of problems require a high demand of computational time, and that is the reason why easily parallelizable methods to solve them have appeared \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress}. %FERGU: TODO reescribir esta frase tan basura

%3 ONE WAY IS TO USE MOEAS BECAUSE
One example of easily parallelizable algorithms to solve MOPs are the Multi-Objective Evolutionary Algorithms (MOEAs), a sub-set of Evolutionary algorithms (EAs). EAs \citep{DBLP:series/ncs/EibenS15DeJong2006} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regards to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met.

 One of the most well-known and referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, and divided into different ranks (based on the
dominance) to be selected for the next generations.

@@

In our preliminary work \citep{Garcia16hpmoon} we tried to demonstrate
% tried to or actually did it? What did we prove? 
that,  instead of making every island focus in a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome in a coevolutionary multi-objective island-based algorithm
can improve the quality of the solutions in the same computing time
when the number of islands increases. We discovered that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. Moreover, the experiments
were conducted in a mono-processor and synchronous island model. This
motivates us to continue this research line; in this paper we propose
a new method that automatically adapts the size of the overlapped
sections. In order to test the validity of the method we have made
tests in a real cluster
with up to 128 nodes, with extra number of islands. % We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ

@@

The hypothesis of this paper is that using overlapped sections of the
chromosome, calculated depending the number of islands to use, in a
coevolutionary multi-objective island-based algorithm can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. %That was the
                                %hypothesis of the previous
                                %paper. This paper has a new
                                %hypothesis, right? - JJ
In order to prove this claim, our new adaptive method has been
compared different overlapping islands schemes have been compared: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
coevolutionary disjoint section approach version of it and a
fixed-size overlapped version. %What was done in the previous paper?
                               %Are we re-using the results in that
                               %paper? - JJ
Different benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.  % Shouldn't this start with "We are trying to
                        % prove..." - JJ

@@

% Parallel multiobjective optimization is the theme. And there are
% papers that do that. Please include them in the state of the art:
% dennis2003parallel (added to bibliography).

@@

%This is an improvement over the previous? What is the relationship?
The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. Their authors concluded that dividing the search space is a good idea, although achieve this is not trivial. Dividing the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separation in processors by objective \citep{Xiao03specialized}. Other authors, such as Martens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative coevolution? - JJ
Addressing high-dimensional problems with cooperative coevolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. This approach has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but using a
master-slave method that splits the population into several processors. As in previous work, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper we have
used up to 128 islands.

@@

\section{Coevolutionary algorithms compared} % This should be called
                                % methodology. The object of the
                                % section is to explain the
                                % methodology used to test the
                                % problem, in order to do that we have
                                % tested different versions that prove
                                % that our methodology works for the
                                % precise reasons we think it works- JJ
\label{sec:coevo}

In this paper we analyse several coevolutionary algorithms we have
implemented by using a parallel multi-objective baseline algorithm (B)
as a basis for comparisons. Our proposed methods %methods or method?
                                %Which methods? - JJ
are based in NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. % This is SoA stuff - JJ
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

%This probably should deserve a "Algorithm" environment too - JJ
This basic distributed algorithms spreads the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different coevolutionary alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
islands. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier. Note that, differently from other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel} all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ

@@

\subsection{Adaptive coevolution with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but
instead using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ
As a first
approach to automatically calculate this value the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were $>$8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Figure \ref{fig:adaptive} explains this method.

>>>>

\begin{abstract}
  % First, state the problem!!!! - JJ
Multi-objective optimization methods have gained interest in the last
years because of the performance they can achieve solving real-world
problems. Many of these problems usually require a high-dimensional
decision space, and multi-objective Evolutionary Algorithms have been
successfully used in the past. Due to the large computational cost
these algorithms require, different parallelization and distribution
methods have been proposed, with different parallel and distributed
computer architectures to cope with them.

One of the most extended
methods are the distributed co-evolutionary algorithms, where each
processor deals only with a part of the decision space, that is, the
individuals belonging to the same sub-population (island) explore the
same subset of decision variables. These subsets are usually disjoint
sets. We have previously proved that extending the length of these
subsets, and therefore, sharing some search space between islands,
implies an improvement in the results.

In this paper, a new method to
automatically adapt the size of the overlapping, or shared, sections
of the individuals has been compared to other collaborative
evolutionary techniques. Different number of islands, chromosome sizes
and problems have been used. The analysis of the obtained experimental
results, by using different metrics, shows that coevolutionary
approaches can provide statistically significant improvements with
respect to the base algorithm, being the method to automatically adapt
the overlapping size the one to obtain the best performance during the
same execution time. Results also show that the relation of the number
of islands (subpopulations) to the length of the chromosome (number of
decision variables) is also a relevant factor to determine the most
efficient alternative to distribute the decision variables. 
\end{abstract}

@@

%2 COMPUTATION -> PARALLEL
These kind of problems require a high demand of computational time, leading to the apparition of easily parallelizable methods to solve them \citep{Luna15Survey,Mukhopadhyay14Survey,Chavez15MO,Hidalgo16residualstress}.
%3 ONE WAY IS TO USE MOEAS BECAUSE
One example of easily parallelizable algorithms to solve MOPs are the Multi-Objective Evolutionary Algorithms (MOEAs), a sub-set of Evolutionary algorithms (EAs). EAs \citep{DBLP:series/ncs/EibenS15} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regards to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met. One of the most well-known and referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, and divided into different ranks (based on the
dominance) to be selected for the next generations.

@@

In our preliminary work \citep{Garcia16hpmoon} we proved
% tried to or actually did it? What did we prove? FERGU: Done
that,  instead of making every island focus in a disjoint subset of the
chromosome, the usage of overlapped (shared) sections of the
chromosome in a coevolutionary multi-objective island-based algorithm
can improve the quality of the solutions in the same computing time
when the number of islands increases. We discovered that the most
adequate overlapping size may vary depending on a relation between the
number of islands and the individual size. However, the experiments
were conducted in a limited experimental system: mono-processor and synchronous island model. This
motivates us to continue this research line using a more complete environment: a real cluster with up to 128 nodes, and a more complete parameter setting. Besides comparing previous methods in this new experimental setup, in this paper we propose
a new method that automatically adapts the size of the overlapped
sections depending on the number of islands to use, comparing it with the previous versions. 
% We have to
                                % emphasize what is actually done. Is
                                % this an addition to the old paper,
                                % or a complete rewrite? - JJ FERGU: rewritten previous paragraph

@@

The hypothesis of this paper is that using overlapped sections of the
chromosome, calculated depending the number of islands to use, in a
coevolutionary multi-objective island-based algorithm, in a real computational environment, can improve the
quality of the solutions in the same computing time with respect to
the baseline, disjoint and partly overlapped methods. %That was the
                                %hypothesis of the previous
                                %paper. This paper has a new
                                %hypothesis, right? - JJ FERGU: no, this is the hypothesis of this paper (comparing adaptive with partly overlapped), but clarifying it
In order to prove this claim, our new adaptive method has been
compared against different overlapping islands schemes: a
baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a
coevolutionary disjoint section approach version of it and a
fixed-size overlapped version. %What was done in the previous paper?
                               %Are we re-using the results in that
                               %paper? - JJ FERGU: explained before that we are showing new results.
Different benchmark problems and different large number of islands
have been used. Results show that this new technique can improve
different quality indicators in the same amount of time when the
problem size is large.  % Shouldn't this start with "We are trying to
                        % prove..." - JJ FERGU: Already writed in previous paragraph

@@

% Parallel multiobjective optimization is the theme. And there are
% papers that do that. Please include them in the state of the art:
% dennis2003parallel (added to bibliography). FERGU: why this paper in particular? It parallelizes the fitness function steps and it is not multi-objective. It does not even describes the GA in detail.

@@

%This is an improvement over the previous? What is the relationship? FERGU: linked
Other kind of approaches have also been explored. The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. Their authors concluded that dividing the search space is a good idea, although achieve this is not trivial. Dividing the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separation in processors by objective \citep{Xiao03specialized}. Other authors, such as Martens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

% Is this really cooperative coevolution? - JJ FERGU: You mean the Dorronsoro one? The title says it
Addressing high-dimensional problems with cooperative coevolution has
also been studied in several works with approaches closer to the one
presented here. The approach to focus on a portion of the chromosome,
as in our overlapped method, was firstly used in the work by
Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining
super-linear performance 
in several instances. This approach has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but using a
master-slave method that splits the population into several processors. As in previous work, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper we have
used up to 128 islands.

@@

\section{Methodology} % This should be called
                                % methodology. The object of the
                                % section is to explain the
                                % methodology used to test the
                                % problem, in order to do that we have
                                % tested different versions that prove
                                % that our methodology works for the
                                % precise reasons we think it works- JJ FERGU: Done, and explained
\label{sec:coevo}

The objective of this section is to explain the methodology we have followed to
compare the different versions of the overlapping schemes.

In this paper we analyse several coevolutionary algorithms we have
implemented by using a parallel multi-objective baseline algorithm (B)
as a basis for comparisons. Our proposed methods %methods or method?
                                %Which methods? - JJ
are based in NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. % This is SoA stuff - JJ FERGU: Yes, I added it here againg because a reviewer wasn't very attentive
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

%This probably should deserve a "Algorithm" environment too - JJ FERGU: You mean the LaTeX one? I've used it in the pseudocode figures
This basic distributed algorithms spreads the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different coevolutionary alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
islands. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

As in the baseline (B), an individual is migrated to another random
island after a fixed number of generations. In the new island, this
individual will be considered as one of the others in the island,
crossed and mutated in the same way, depending on the island
identifier. Note that, differently from other works such as the ones
described by Talbi et al. \citep{Talbi08Parallel} all the islands deal
with complete chromosomes for fitness calculation, so our approach can
deal with decomposable and no decomposable problems. %Don't understand what you say
                                %here. What you are saying is that our
                                %approach can't deal with decomposable
                                %problems, actually. - JJ FERGU: you are right, fixed

@@

\subsection{Adaptive coevolution with overlapping islands (A)} 
As in the previous method, the sections to address are overlapped, but
instead using one extra section on each side of the $p$ section ($p+1$
and $p-1$), it uses $c$ fragments on each side ($p+c$ and $p-c$),
being $c$ a value depending on the number of islands.
% That's not adaptive. It's simply variable - JJ FERGU: it depends on the number of islands. What term do you suggest?
As a first
approach to automatically calculate this value the results shown in
\citep{Garcia16hpmoon} have been used as a base to obtain $c$. In that
work, the overlapped method ($c=1$) obtained better results when the
number of islands were $>$8. On the contrary, when the number of
islands is small, no overlapping is necessary. So, we have used this
knowledge to create the formula $c=round(0.2*P-1)/2$ to calculate the
extra sections to overlap. Therefore, when $P=8$ then $c=0$
(equivalent to D), when $P=16$ then $c=1$ (equivalent to O), and so
on. Figure \ref{fig:adaptive} explains this method.
